---
title: "Tech Dividends, Part 1"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()      
---

```{r, message=FALSE, warning = FALSE}
library(data.table)
library(magrittr)
library(stringr)
library(rlang)
library(ggplot2)
library(riingo)
library(tidyquant)
library(kableExtra)
library(RplotterPkg)
library(RtsaPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  The following notes/R scripts were inspired from an article [Tech Dividends, Part 1](https://rviews.rstudio.com/2019/08/07/tech-dividends-part1/) by Jonathan Regenstein.
</div>  

<div class="note">Note: Be aware that this application has an internet request for dividend data that will take time to complete.</div>

# Tech Dividends, Part 1
## The data
### Set the Tiingo API key token
```{r}
riingo::riingo_set_token(RtsaPkg::read_tingo_key())
```

### Dividend prices for companies over a 10 year period
1. Get the companies listed on the NASDAQ:
```{r}
nasdaq_dt <- data.table::setDT(tidyquant::tq_exchange("NASDAQ")) 
str(nasdaq_dt)
```

2. Rename variables:
```{r}
nasdaq_new_dt <- nasdaq_dt[, .(Symbol = symbol, Company = company, LastSalePrice = last.sale.price, MarketCap = market.cap, IpoYear = ipo.year, Sector = sector)]
```
3. Make *MarketCap* numeric:
```{r}
marketcap <- nasdaq_new_dt$MarketCap
marketcap <- str_remove_all(marketcap, "\\$")
mil <- 1000000
bil <- 1000*mil

mils <- as.numeric(str_replace_all(marketcap,"[M]",""))*mil
bils <- as.numeric(str_replace_all(marketcap,"[B]",""))*bil
check_val <- function(i){
  if(is.na(mils[[i]])){
    return(bils[[i]]/1000)
  }else {
    return(mils[[i]]/1000)
  }
}
vals_lst <- lapply(1:length(mils), check_val)
nasdaq_new_dt <- nasdaq_new_dt[, MarketCap := unlist(vals_lst)]
data.table::setorderv(nasdaq_new_dt, cols = "MarketCap", order = -1, na.last = T)
```

4. Get the NASDAQ company symbols based on *MarketCap* and that have IPO'd since 2007:
```{r}
nasdaq_new_dt <- nasdaq_new_dt[IpoYear > 2007,.(Symbol, MarketCap)]
nasdaq_new_dt[1:10]
```
<div class="note">Note: There are 1484 companies that have IPO'd since 2007</div>

5. Get the dividends posted among the top 600 NASDAQ companies:
```{r}
nasdaq_dividends_dt <-  RtsaPkg::get_tiingo_dividends(symbols = nasdaq_new_dt$Symbol[1:600], from = "2009-01-01", to = "2019-12-31")
str(nasdaq_dividends_dt)
```

6. Add a *year* variable:
```{r}
nasdaq_dividends_dt[, Year := as.numeric(format(Date, "%Y"))]
```
7. Select variables:
```{r}
nasdaq_dividends_dt <- nasdaq_dividends_dt[, .(Symbol,Date,Year,Close,DivCash)]
str(nasdaq_dividends_dt)
```

## First date when the companies posted a dividend
1. Get the first observation from each *Symbol* grouping where a dividend was posted:
```{r}
first_dividend_dt <- nasdaq_dividends_dt[DivCash > 0, .SD[1L], by = Symbol]
kbl(
  x = first_dividend_dt,
  format = "html",
  caption = "First Posted Dividend Dates for Top NASDAQ Companies",
  align = c("c", "c", "c", "c", "c")) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```

## Count for each year when a company first granted a dividend
```{r}
first_dividends_year_dt <- first_dividend_dt[, .N, by = Year][order(Year)]
kbl(
  x = first_dividends_year_dt,
  format = "html",
  caption = "Counts of First Posted Dividend Dates by Year for Top NASDAQ Companies",
  align = c("c", "c")) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```

## Most recent posting in each company.
1. Get the most recent dividend posting from each company:
```{r}
latest_dividends_dt <- nasdaq_dividends_dt[DivCash > 0, .SD[.N], by = Symbol][order(-DivCash)]
```
<div class="note">Note: The company "ZG" had a very large value and is filtered out in the plot below. </div>

2. Plot the most recent dividends posted:
```{r,fig.width=12}
RplotterPkg::create_scatter_plot(
  df = latest_dividends_dt,
  aes_x = "Date",
  aes_y = "DivCash",
  title = "Last Dividends Paid by NASDAQ Companies",
  x_title = "Month",
  y_title = "Dividends($)",
  rot_y_tic_label = TRUE,
  x_limits = c(as.Date("2013-12-31"), as.Date("2020-01-01")),
  x_major_date_breaks = "6 month",
  x_date_labels = "%b-%Y",
  y_limits = c(0.0, 5.0),
  y_major_breaks = seq(0.0,5.0,0.5),
  pts_size = 1.2,
  pts_color = "blue",
  pts_fill = "blue",
  show_minor_grids = F
) + ggrepel::geom_text_repel(mapping = aes(label = Symbol))
```
3. A more focused plot:
```{r,fig.width=12}
RplotterPkg::create_scatter_plot(
  df = latest_dividends_dt,
  aes_x = "Date",
  aes_y = "DivCash",
  title = "Last Dividends Paid by NASDAQ Companies",
  x_title = "Month",
  y_title = "Dividends($)",
  rot_y_tic_label = TRUE,
  x_limits = c(as.Date("2018-12-31"), as.Date("2020-01-01")),
  x_major_date_breaks = "1 month",
  x_date_labels = "%b-%Y",
  y_limits = c(0.0, 0.7),
  y_major_breaks = seq(0.0,0.7,0.1),
  pts_size = 1.2,
  pts_color = "blue",
  pts_fill = "blue",
  show_minor_grids = F
) + ggrepel::geom_text_repel(mapping = aes(label = Symbol))
```

## Plot all the dividends posted by each company
```{r, fig.height=10, fig.width=12}
RplotterPkg::create_scatter_plot(
  df = nasdaq_dividends_dt,
  aes_x = "Date",
  aes_y = "DivCash",
  aes_color = "Symbol",
  title = "All Dividends Paid by NASDAQ Companies",
  x_title = "Year",
  y_title = "Dividends($)",
  rot_y_tic_label = TRUE,
  x_limits = c(as.Date("2013-12-31"), as.Date("2020-01-01")),
  x_major_date_breaks = "1 year",
  x_date_labels = "%Y",
  y_limits = c(0.0, 5.0),
  y_major_breaks = seq(0.0,5.0,0.5),
  pts_size = 1.2,
  pts_color = "blue",
  pts_fill = "blue",
  show_minor_grids = F,
  show_legend = F
) 
```

## Grouped statistical summaries
### Top 10 companies' dividend stats for 2019
```{r}
stats_comp_dividends_dt <- nasdaq_dividends_dt[Date > as.Date("2018-12-31"), .(N = .N, Mean = mean(DivCash), Median = median(DivCash), Min = min(DivCash), Max = max(DivCash), Total = sum(DivCash)), by = Symbol][order(-Total)]
kbl(
  x = stats_comp_dividends_dt[1:10],
  format = "html",
  caption = "Dividend Statistics for the Top 10 NASDAQ Companies in 2019",
  align = c("c", "c", "c", "c", "c", "c")) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```

### Dividend stats by year
```{r}
stats_year_dividends_dt <- nasdaq_dividends_dt[, .(N = .N, Mean = mean(DivCash), Median = median(DivCash), Min = min(DivCash), Max = max(DivCash), Total = sum(DivCash)), by = Year][order(Year)]
kbl(
  x = stats_year_dividends_dt,
  format = "html",
  caption = "Dividend Statistics across the years for NASDAQ Companies",
  align = c("c", "c", "c", "c", "c", "c")) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```
