---
title: "Tech Dividends, Part 2"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()     
---

```{r, message=FALSE, warning = FALSE}
library(data.table)
library(magrittr)
library(ggplot2)
library(ggrepel)
library(RplotterPkg)
library(RtsaPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  The following notes/R scripts were inspired from an article [Tech Dividends, Part 2](https://rviews.rstudio.com/2019/08/17/tech-dividends-part-2/) by Jonathan Regenstein.
</div>  


# Tech Dividends, Part 2
## The data
### Set the Tiingo API key token
```{r}
riingo::riingo_set_token(RtsaPkg::read_tingo_key())
```

## History of posted dividends
1. Get 10 years of historic data between the dates 2009-01-01 and 2019-12-31:
```{r}
symbols <- c("IBM","HPQ","TXN","CSCO","INTC","ORCL","AAPL","MSFT","QCOM")

dividends_dt <- RtsaPkg::get_tiingo_dividends(symbols = symbols, from = "2009-01-01", to = "2019-12-31")
dividends_dt <- dividends_dt[, .(Symbol, Date, DivCash, Close)]
str(dividends_dt)
```
2. Plot the historic data:
```{r,fig.width=12, fig.height=12}
RplotterPkg::create_scatter_plot(
  df = dividends_dt,
  aes_x = "Date",
  aes_y = "DivCash",
  aes_color = "Symbol",
  title = "Dividends Posted by 9 Attractive Companies",
  subtitle = "2009-01-01 to 2019-12-31",
  x_title = "Year",
  y_title = "Dividends($)",
  x_limits = as.Date(c("2009-01-01", "2020-01-01")),
  x_major_breaks = seq(as.Date("2009-01-01"), as.Date("2020-01-01"), by="1 year"),
  x_date_labels = "%Y",
  rot_y_tic_label = TRUE,
  y_limits = c(0.0, 2.0),
  y_major_breaks = seq(0.0,2.0,0.2),
  pts_size = 1.2,
  show_minor_grids = FALSE
)
```

## Most recent dividend
1. Get the most recent dividend posted:
```{r}
most_recent_df <- dividends_dt[, .SD[.N], by = Symbol]
```

2. Plot the latest postings:
```{r}
RplotterPkg::create_scatter_plot(
  df = most_recent_df,
  aes_x = "Date",
  aes_y = "DivCash",
  aes_color = "Symbol",
  title = "Last Dividend Posted by 9 Attractive Companies in 2019",
  x_title = "December",
  y_title = "Dividends($)",
  x_limits = as.Date(c("2019-12-01", "2020-01-01")),
  x_major_breaks = seq(as.Date("2019-12-01"), as.Date("2020-01-01"), by="1 day"),
  x_date_labels = "%d",
  rot_y_tic_label = TRUE,
  y_limits = c(0.0, 0.7),
  y_major_breaks = seq(0.0,0.7,0.1),
  pts_size = 1.2,
  show_minor_grids = FALSE,
) + ggrepel::geom_text_repel(mapping = aes(label = Symbol))
```

## Total Annual Yield for 2019

<div class="note">Note: The Total Annual Yield (TAY) is defined as the sum of total dividends in 2019 divided by the closing price on the first dividend date.</div>

1. Find the first dividend for each company in 2019:
```{r}
first_dividend_2019_dt <- dividends_dt[DivCash > 0 & Date > as.IDate("2018-12-31"), .SD[1L], by = Symbol][order(-DivCash)]
first_dividend_2019_dt[1:10]
```
2. Find the sum of the total dividends in 2019:
```{r}
sum_dividend_2019_dt <- dividends_dt[DivCash > 0 & Date > as.IDate("2018-12-31"), lapply(.SD, sum), by = Symbol, .SDcols = "DivCash"][order(-DivCash)]
sum_dividend_2019_dt <- sum_dividend_2019_dt[, .(Symbol, TotalDivCash = DivCash)]
sum_dividend_2019_dt[1:10]
```
3. Outer join both data.tables:
```{r}
data.table::setkey(first_dividend_2019_dt, "Symbol")
data.table::setkey(sum_dividend_2019_dt, "Symbol")
tay_dividend_2019_dt <- first_dividend_2019_dt[sum_dividend_2019_dt]
tay_dividend_2019_dt[1:10]
```

4. Compute `TAY = (TotalDivCash/Close) * 100`:
```{r}
tay_dividend_2019_dt <- tay_dividend_2019_dt[, TAY := (TotalDivCash/Close)*100][order(-TAY)]
tay_dividend_2019_dt[1:10]
```

2. Plot the TAY for the companies as a bar chart:
```{r}
RplotterPkg::create_bar_plot(
  df = tay_dividend_2019_dt,
  aes_x = "Symbol",
  aes_y = "TAY",
  title = "Total Annual Yield for 2019",
  x_title = "Symbol",
  y_title = "Yield (%)",
  do_coord_flip = T,
  order_bars = "asc",
  rot_y_tic_label = T,
  bar_fill = "green",
  bar_alpha = 0.5
)
```
