---
title: "Dividend Discovery"
output: 
   html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()       
---

```{r, message=FALSE, warning = FALSE}
library(data.table)
library(rlang)
library(ggplot2)
library(ggrepel)
library(riingo)
library(tidyquant)
library(kableExtra)
library(RplotterPkg)
library(RtsaPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.width = 12, fig.height = 8)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  The following notes/R scripts were inspired from an article [Dividend Discovery](http://www.reproduciblefinance.com/2019/07/10/dividend-discovery/) by Jonathan Regenstein.
</div>  

<div class="note">Note: Be aware that this application has an internet request for dividend data that will take time to complete.</div>

# Dividend Discovery
## The data

### Set the Tiingo API key token
```{r}
riingo::riingo_set_token(RtsaPkg::read_tingo_key())
```

### Prepare the S&P 500 data
1. Using `tidyquant::tq_index()` get the tickers for S&P 500 index.
```{r}
sp_500_dt <- data.table::setDT(tidyquant::tq_index("SP500"))
str(sp_500_dt)
```
2. Order the tickers by weight and select the top 30 S&P 500 tickers.
```{r}
tickers_dt <- sp_500_dt[order(-weight)][1:30]
tickers_dt[1:10]
```
3. Get historic stock market prices for the top 30 S&P 500 between "1990-01-01" and "2018-12-31":
```{r}
prices_dt <- RtsaPkg::get_tiingo_stock_prices(symbols = tickers_dt$symbol, from = "1990-01-01", to = "2018-12-31")
str(prices_dt)
```
4. Select *Symbol*, *Date*, *Close*, *DivCash*
```{r}
dividends_dt <- prices_dt[, .(Date, Close, DivCash, Date_alt = as.IDate(Date)), by = Symbol]
str(dividends_dt)
```
<div class="note">Note: There are 29 symbols. One ticker is not supported by Tiingo -- "BRK.B" Berkshire Hathaway Inc. Class B</div>

## Dividend payments
### Select the most recent dividend payment
```{r}
dividends_latest_dt <- dividends_dt[DivCash > 0, .SD[.N], by = Symbol][order(-DivCash)]
dividends_latest_dt[1:10]
```


<div class="note">Note: the stock split that happened in 2014 for Google paid in the form of a dividend.  We will filter this amount out along with just last dividends from 2018.</div>

### Get the last posted dividends in 2018
```{r}
dividends_last_2018_dt <- dividends_dt[DivCash > 0 & DivCash < 200 & Date_alt > as.IDate("2017-12-31"),.SD[.N], by = Symbol][order(-DivCash)]
str(dividends_last_2018_dt)
```
<div class="note">Note: we have 21 companies that posted dividends in the year 2018.</div>

### Plot the dividends last posted in 2018
```{r, fig.width=12}
RplotterPkg::create_scatter_plot(
  df = dividends_last_2018_dt,
  aes_x = "Date_alt",
  aes_y = "DivCash",
  title = "Last Dividends Paid by S&P 500 Companies",
  subtitle = "Top 21 in year 2018",
  x_title = "Month",
  y_title = "Dividends($)",
  x_limits = as.Date(c("2018-10-01", "2018-12-31")),
  x_major_breaks = seq(as.Date("2018-10-01"), as.Date("2018-12-31"), by="1 week"),
  x_date_labels = "%b %d",
  rot_y_tic_label = T,
  y_limits = c(0.0, 2.0),
  y_major_breaks = seq(0.0,2.0,0.2),
  pts_size = 1.2,
  pts_color = "blue",
  pts_fill = "blue",
  show_minor_grids = F
) + ggrepel::geom_text_repel(mapping = aes(label = Symbol))
```
  
## Compute the Total Annual Yield for 2018 
<div class="note">Note: The Total Annual Yield (TAY) is defined as the sum of total dividends in 2018 divided by the closing price on the first dividend date.</div>
1. Find the first dividend for each company in 2018:
```{r}
first_dividend_2018_dt <- dividends_dt[DivCash > 0 & DivCash < 200 & Date_alt > as.IDate("2017-12-31"), .SD[1L], by = Symbol][order(-DivCash)]
first_dividend_2018_dt[1:10]
```
2. Find the sum of the total dividends in 2018:
```{r}
sum_dividend_2018_dt <- dividends_dt[DivCash > 0 & DivCash < 200 & Date_alt > as.IDate("2017-12-31"), lapply(.SD, sum), by = Symbol, .SDcols = "DivCash"][order(-DivCash)]
sum_dividend_2018_dt <- sum_dividend_2018_dt[, .(Symbol, TotalDivCash = DivCash)]
sum_dividend_2018_dt[1:10]
```
3. Outer join both data.tables:
```{r}
data.table::setkey(first_dividend_2018_dt, "Symbol")
data.table::setkey(sum_dividend_2018_dt, "Symbol")
tay_dividend_2018_dt <- first_dividend_2018_dt[sum_dividend_2018_dt]
tay_dividend_2018_dt[1:10]
```
4. Compute `TAY = (TotalDivCash/Close) * 100`:
```{r}
tay_dividend_2018_dt <- tay_dividend_2018_dt[, TAY := (TotalDivCash/Close)*100][order(-TAY)]
tay_dividend_2018_dt[1:10]
```
5. Plot the TAY:
```{r}
RplotterPkg::create_scatter_plot(
  df = tay_dividend_2018_dt,
  aes_x = "Date_alt",
  aes_y = "TAY",
  title = "Total Annual Yield(TAY) of Dividends",
  subtitle = "Top companies in year 2018",
  x_title = "Month",
  y_title = "Yield(%)",
  x_limits = as.Date(c("2018-01-01", "2018-08-01")),
  x_major_breaks = seq(as.Date("2018-01-01"), as.Date("2018-08-01"), by="1 month"),
  x_date_labels = "%b",
  rot_y_tic_label = TRUE,
  y_limits = c(0.0, 6.0),
  y_major_breaks = seq(0.0,6.0,0.5),
  pts_size = 1.2,
  pts_color = "blue",
  pts_fill = "blue",
  show_minor_grids = F
) + ggrepel::geom_text_repel(mapping = aes(label = Symbol))
```

## Dividends across the years
1. Add a *year* variable to `dividends_dt`:
```{r}
dividends_dt[, Year := as.numeric(format(Date_alt, "%Y"))]
dividends_dt[1:10]
```

2. Compute the total yearly dividend payout for each company each year:
```{r}
dividends_yearly_dt <- dividends_dt[, .(Total_Div = sum(DivCash)), by = .(Symbol, Year)]
dividends_yearly_dt[1:10]
```

<div class="note">Note: For *AAPL* there was a stock split in years 2012, 2013, 2014 and then back down to $2 in 2015.</div>

3. Create a column that shows increase/decrease from the previous year:
```{r}
dividends_yearly_dt[, Div_Increase := ifelse(Total_Div > shift(Total_Div,1), 1, 0)]
dividends_yearly_dt <- na.omit(dividends_yearly_dt, cols = "Div_Increase")
dividends_yearly_dt[1:10]
```

4. Get the total increases for each company:
```{r}
dividend_increases_dt = dividends_yearly_dt[, lapply(.SD, sum), by = Symbol, .SDcols = "Div_Increase"]
dividend_increases_totals_dt <- dividend_increases_dt[, .(Symbol, Div_Increase_Total = Div_Increase)]
dividend_increases_totals_dt <- dividend_increases_totals_dt[order(-Div_Increase_Total)]
dividend_increases_totals_dt[1:10]
```

5. Plot the increases:
```{r}
RplotterPkg::create_bar_plot(
  df = dividend_increases_totals_dt,
  aes_x = "Symbol",
  aes_y = "Div_Increase_Total",
  do_coord_flip = T,
  rot_y_tic_label = T,
  y_title = "Years of Increase",
  order_bars = "asc"
)
```
