---
title: "Looking Back on 2018"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    css: ../style.css
params:
  date: !r Sys.Date()       
---

```{r, message=FALSE, warning = FALSE}
library(data.table)
library(ggplot2)
library(riingo)
library(kableExtra)
library(magrittr)
library(grid)
library(gtable)
library(RplotterPkg)
library(RtsaPkg)
```

```{r,setup, include=FALSE, eval=TRUE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message=FALSE, fig.width = 12, fig.height = 8)
```

<div>Author: Rick Dean</div>
<div>Article date: `r params$date`</div>

<div class="abstract">
  <p class="abstract">Abstract</p>
  The following notes/R scripts were inspired from an article [Looking Back on 2018](http://www.reproduciblefinance.com/2019/01/14/looking-back-on-last-year/) by Jonathan Regenstein.
</div>  

# Looking Back on 2018
## The data
### Set the Tiingo API key token
```{r}
riingo::riingo_set_token(RtsaPkg::read_tingo_key())
```

### Get the daily prices

1. Set 10 sector Exchange-Traded Funds (ETF's i.e. sector indexes) plus the SP500 index (SPY) along with their corresponding sector labels.
```{r}
symbol_sector_dt <- data.table::data.table(
  Symbol = c("XLY", "XLP", "XLE",
          "XLF", "XLV", "XLI", "XLB",
          "XLK", "XLU", "XLRE",
          "SPY"),
  Sector = c("Consumer Discretionary", "Consumer Staples", "Energy",
          "Financials", "Health Care", "Industrials", "Materials",
          "Information Technology", "Utilities", "Real Estate",
          "Market")
)
```
2. Get the daily prices for 2018 for the symbols:
```{r, message = FALSE, warning = FALSE}
prices_dt <- RtsaPkg::get_tiingo_stock_prices(symbols = symbol_sector_dt$Symbol, from = "2017-12-29", to = "2018-12-31")
prices_dt <- prices_dt[, .(Symbol, Date, AdjClose)]
str(prices_dt)
```
3. Outer join both `ticker_sector_dt` and `prices_dt` using *Symbol* as the key:
```{r}
data.table::setkey(symbol_sector_dt, "Symbol")
data.table::setkey(prices_dt, "Symbol")
prices_dt <- prices_dt[symbol_sector_dt]
prices_dt[, Sector := as.factor(Sector)]
prices_dt[1:10]
```

## Group by sector and show the first index
```{r}
first_index_dt <- prices_dt[, .SD[1], by = Sector]
```
```{r}
first_index_dt %>% 
kbl(
  format = "html",
  caption = "First Prices Across Sectors",
  align = c("c", "c", "c", "c"),
  digits = 1) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```

<div class="note">Note: All showing the starting date of 2017-12-29</div>

## Group by Sector and compute the daily return

The "DailyReturn" column is the log10 difference between the current adjusted value and its prior daily adjusted value.
```{r}
prices_daily_return_dt <- prices_dt[, DailyReturn := log10(AdjClose) - log10(shift(AdjClose, 1)), by = Sector]
prices_daily_return_dt <- na.omit(prices_dt, cols = "DailyReturn")
str(prices_daily_return_dt)
```
<div class="note">Note: the use of *shift()* and that the number of observations goes from 2772 to 2761 (11 sector observations from 2017-12-29 were lost from computing the first shift).
Also our dates are starting at "2018-01-02".

### Tabulate the first row of each sector group (2018-01-02)
```{r}
prices_daily_return_dt[, .SD[1], by = Sector] %>% 
  kbl(
  format = "html",
  caption = "First DailyReturn Across Sectors",
  align = c("c", "c", "c", "c", "c")) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```

### Tabulate the second row of each sector group (2018-01-03)
```{r}
prices_daily_return_dt[, .SD[2], by = Sector] %>% 
  kbl(
  format = "html",
  caption = "Second DailyReturn Across Sectors",
  align = c("c", "c", "c", "c", "c")) %>% 
kable_paper(full_width = T, position = "center") %>% 
row_spec(0, background = "purple", color = "white, font_size = 22")
```

## Plot the *DailyReturn* across time for each sector group
```{r, fig.height=14}
RplotterPkg::multi_stick_plot(
  df = prices_daily_return_dt,
  factor_var = "Sector",
  factor_x = "Date",
  aes_y = "DailyReturn",
  title = "Daily Returns for Stock Sectors",
  subtitle = "Jan 2018 to Jan 2019",
  x_title = "Month",
  columns = 3,
  rot_y_tic_label = T,
  x_limits = as.Date(c("2018-01-01", "2019-01-01")),
  x_major_breaks = seq(as.Date("2018-01-01"), as.Date("2019-01-01"), by="2 month"),
  x_date_labels = "%b",
  y_limits = c(-0.03, 0.06),
  y_major_breaks = seq(-0.03, 0.06, 0.03),
  row_height = 3.0
)
```

## Plot positive and negative prices
1. Create two numeric variables *ReturnPos* and *ReturnNeg*:
```{r}
prices_pos_neg_dt <- prices_daily_return_dt[, .(Sector, Date, ReturnPos = ifelse(DailyReturn > 0, DailyReturn, 0.0), ReturnNeg = ifelse(DailyReturn < 0, DailyReturn, 0.0))]
str(prices_pos_neg_dt)
```
2. Create a long dataframe involving variables "return_pos" and "return_neg".
```{r}
pos_neg_long_dt <- data.table::melt(prices_pos_neg_dt, 
                                    id.vars = c("Sector","Date"), 
                                    measure.vars = c("ReturnPos", "ReturnNeg"), 
                                    variable.name = "Direction", value.name = "DailyReturn")
str(pos_neg_long_dt)
```
3. Plot the two directions across sectors:
```{r, fig.height=14}
RplotterPkg::multi_stick_plot(
  df = pos_neg_long_dt,
  factor_var = "Sector",
  factor_x = "Date",
  title = "2018 Net Daily Returns",
  columns = 3,
  aes_y = "DailyReturn",
  aes_color = "Direction",
  rot_y_tic_label = T,
  x_limits = as.Date(c("2018-01-01", "2019-01-01")),
  x_major_breaks = seq(as.Date("2018-01-01"), as.Date("2019-01-01"), by="2 month"),
  x_date_labels = "%b",
  y_limits = c(-0.03, 0.06),
  y_major_breaks = seq(-0.03, 0.06, 0.03),
  row_height = 3.0,
  show_legend = F
)
```

## Plot Information Technology

Plot just the "Information Technology" returns with the major time tics showing the location of each quarter.

1. Filter out "Information Technology":
```{r}
info_tech_dt <- pos_neg_long_dt[Sector == "Information Technology"]
```
2. Plot the positive and negative:
```{r, fig.width=12, fig.height=10}
RplotterPkg::create_stick_plot(
  df = info_tech_dt,
  aes_x = "Date",
  aes_y = "DailyReturn",
  aes_color = "Direction",
  title = "2018 Information Tech Net Daily Returns",
  subtitle = "Jan 2018 - Jan 2019",
  rot_y_tic_label = T,
  x_limits = as.Date(c("2018-01-01", "2019-01-01")),
  x_major_breaks = seq(as.Date("2018-01-01"), as.Date("2019-01-01"), by="2 month"),
  x_date_labels = "%b",
  y_limits = c(-0.03, 0.06),
  y_major_breaks = seq(-0.03, 0.06, 0.03),
  show_legend = F
)
```
